A Standard for the Scholarly Citation of Archaeological Data

### 2017-08-27 analysis/paper/paper.Rmd
guides and best practices

@@

@@

@@

Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S., ... & Errington, T. M. (2016). Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency. PLoS Biol, 14(5), e1002456.

@@

@@

@@

@@

>>>>

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).

Several guides with detailed information relevant to preparing archaeological data for sharing on a repository.

- http://guides.archaeologydataservice.ac.uk/
- http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction

- http://www.icpsr.umich.edu/files/ICPSR/access/dataprep.pdf
- https://www.dataone.org/sites/all/documents/DataONE_BP_Primer_020212.pdf
- http://www.data-archive.ac.uk/media/2894/managingsharing.pdf

@@

In-text citation:

“This paper uses data from the [name] data repository at https://doi.org/***** (Jones et al. 2008a), first described in Jones et al. 2008b. “

Data reference and article reference in reference list:

Jones A, Bloggs B, Smith C (2008a). <Title of data package>. <Repository name>. doi: https://doi.org/#####. [Version and/or date of access].

@@

Cranston, K., Harmon, L. J., O'Leary, M. A., & Lisle, C. (2014). Best practices for data sharing in phylogenetic research. PLoS currents, 6.doi: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645

@@

Kansa, E. C., Kansa, S. W., & Watrall, E. (2011). Archaeology 2.0: new approaches to communication and collaboration. Cotsen Digital Archaeology series. http://escholarship.org/uc/item/1r6137tb

Kansa, E. C., & Kansa, S. W. (2013). Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication. Journal of Eastern Mediterranean Archaeology & Heritage Studies, 1(1), 88–97. https://doi.org/10.5325/jeasmedarcherstu.1.1.0088

Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S. & Errington, T. M. (2016). Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency. PLoS Biol, 14(5), e1002456.

Kintigh, K. (2006). The Promise and Challenge of Archaeological Data Integration. American Antiquity, 71(3), 567–578. https://doi.org/10.1017/S0002731600039810

@@

Mounce, R. (2014). Open data and palaeontology. In  Moore, S. (ed.) _Issues in open research data_, Ubiquity Press, London. DOI: https://doi.org/10.5334/ban.j

@@

Penev, L., Mietchen, D., Chavan, V., Hagedorn, G., Smith, V., Shotton, D. & Groom, Q. (2017). Strategies and guidelines for scholarly publishing of biodiversity data. Research Ideas and Outcomes, 3, e12431.

@@

Thessen, A. E., & Patterson, D. J. (2011). Data issues in the life sciences. ZooKeys, (150), 15.

@@

Whitlock, M. C. (2011). Data archiving in ecology and evolution: best practices. Trends in Ecology & Evolution, 26(2), 61-65.
